---
title: "Application to simple datasets"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Application to simple datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(grouper)
library(ompr)
library(ompr.roi)
#library(ROI.plugin.gurobi)
library(ROI.plugin.glpk)
```

# Introduction 

This vignette illustrates the use of the package on simple datasets, for which
the optimal solutions are apparent from inspection.

# Diversity-Based Assignment

## Dataset 001 (diversity only)

The first dataset comprises just 4 students. Here is what it looks like. The
name of this dataset indicates that it is for the diversity-based-assignment
(dba) model and that it consists of the group composition (gc) information.

```{r}
dba_gc_ex001
```

It is intuitive that an assignment into two groups of size two, based on the
diversity of majors alone, should assign students 1 and 2 into the first 
group and the remaining two students into another group.

The corresponding YAML `dba_gc_ex001.yml` file for this exercise consists of the
following lines:

```{r echo=FALSE, comment=''}
cat(readLines(system.file("extdata",  "dba_params_ex001.yml",   
                          package = "grouper")),  sep = '\n')
```

To run the assignment using only the primary major (ignoring the skill), we 
can use the following commands. We can use either the gurobi solver, or the 
glpk solver for this example. Both are equally fast.

```{r}
# indicate appropriate columns using integer ids.
df_ex001_list <- extract_student_info(dba_gc_ex001, "diversity",
                                  demographic_cols = 2, skills = 3, 
                                  self_formed_groups = 4)
yaml_ex001_list <- extract_params_yaml(system.file("extdata", 
                                             "dba_params_ex001.yml",  
                                             package = "grouper"),
                                       "diversity")
m1 <- prepare_model(df_ex001_list, yaml_ex001_list, assignment="diversity",
                    w1=1.0, w2=0.0)
#result3 <- solve_model(m1, with_ROI(solver="gurobi"))
result3 <- solve_model(m1, with_ROI(solver="glpk"))
assign_groups(result3, assignment = "diversity", dframe=dba_gc_ex001, 
              group_names="groups")
```

We can see that students 1 and 2 have been assigned to topic 1, repetition 1. Students 
3 and 4 have been assigned to topic 2, repetition 1.

## Dataset 001 (skills only)

```{r}
# indicate appropriate columns using integer ids.
df_ex001_list <- extract_student_info(dba_gc_ex001, "diversity",
                                  demographic_cols = 2, skills = 3, 
                                  self_formed_groups = 4)
yaml_ex001_list <- extract_params_yaml(system.file("extdata", 
                                             "dba_params_ex001.yml",  
                                             package = "grouper"),
                                       "diversity")
m1a <- prepare_model(df_ex001_list, yaml_ex001_list, assignment="diversity",
                    w1=0.0, w2=1.0)
#result3 <- solve_model(m1a, with_ROI(solver="gurobi"))
result3 <- solve_model(m1a, with_ROI(solver="glpk"))

assign_groups(result3, assignment = "diversity", dframe=dba_gc_ex001, 
              group_names="groups")

get_solution(result3, smin)
get_solution(result3, smax)
```

We can see that students 1 and 2 have been assigned to topic 1, repetition 1. Students 
3 and 4 have been assigned to topic 2, repetition 1.

## Dataset 003

This dataset demonstrates the use of a custom dissimilarity matrix instead of
using the default Gower distance from the
[cluster](https://cran.r-project.org/package=cluster) package.

```{r}
dba_gc_ex003
```

Now consider a situation where we wish to consider years 1 and 2 different from
years 3 and 4, and `math` and `dsds` (STEM majors) to be different from `elts`
and `history` (non-STEM majors). For each difference, we assign a score of 1.

This means that students 1 and 2 would have a dissimilarity score of 1 due to 
their difference in majors. Students 1 and 3 would also have a score of 1, but 
due to their difference in years. Students 1 and 4 would have score of 2, due to
their differences in majors and in years. The overall dissimilarity matrix would 
be:

```{r}
d_mat <- matrix(c(0, 1, 1, 2,
                  1, 0, 2, 1,
                  1, 2, 0, 1,
                  2, 1, 1, 0), nrow=4, byrow = TRUE)
```

To run the optimisation for this model, we can execute the following code:

```{r}
df_ex003_list <- extract_student_info(dba_gc_ex003, "diversity",
                                skills = NULL,
                                self_formed_groups = 3,
                                d_mat=d_mat)
yaml_ex003_list <- extract_params_yaml(system.file("extdata",   
                                                   "dba_params_ex003.yml",   
                                                   package = "grouper"), 
                                       "diversity")
m3 <- prepare_model(df_ex003_list, yaml_ex003_list, w1=1.0, w2=0.0)
result <- solve_model(m3, with_ROI(solver="glpk", verbose=TRUE))

assign_groups(result, "diversity", dba_gc_ex003, group_names="self_groups")
```

As you can see, the members of the two groups have maximal difference between 
them - they differ in terms of their year, and in terms of their major.

## Dataset 004

In this example, we demonstrate that `grouper` provides the flexibility to constrain
group sizes for individual topics. This could be useful in a situation where a 
particular project topic may require a larger group.

The dataset we use contains only skill levels (Python skills, higher corresponding
to more skill).

```{r}
dba_gc_ex004
```

Suppose we wish to assign the students to two topics, but the second topic requires
3 members, and the first requires only 2. In this example, we only utilise the 
skill levels; no demographic variables are included in the objective function.

```{r}
df_ex004_list <- extract_student_info(dba_gc_ex004, 
                                      skills = 2, 
                                      self_formed_groups = 3, 
                                      d_mat=matrix(0, 5, 5))
yaml_ex004_list <- extract_params_yaml(system.file("extdata",    
                                                   "dba_params_ex004.yml",   
                                                   package = "grouper"),  
                                       "diversity")
m4 <- prepare_model(df_ex004_list, yaml_ex004_list, w1=0.0, w2=1.0)
result <- solve_model(m4, with_ROI(solver="glpk", verbose=TRUE))
assign_groups(result, "diversity", dba_gc_ex004, group_names="self_groups")
```

Due to the constraints, topic 2 was assigned 3 members, while preserving the 
total skill level in each group (to be 4).

# Preference-Based Assignment

## Dataset 002

The second datasets comprises 8 students. Here is a listing of the dataset:

```{r}
pba_gc_ex002
```

Each student is in a self-formed group of size 2, indicated via the `grouping`
column. Suppose that, for this set of students, the instructor wishes to assign
students into two topics, with each topic having two sub-groups. This requires
the preference matrix to have 4 columns - one for each topic-subgroup
combination. Remember that the ordering of topics/subtopics should be:

T1S1, T2S1, T1S2, T2S2

There should be 4 rows in the preference matrix - one for each self-formed
group.

```{r}
pba_prefmat_ex002
```

It is possible to assign each self-formed group to its optimal choice of
topic-subtopic combination. In our solution, we should see that group 1 is
assigned to subtopic 1 of topic 1, group 2 is assigned to sub-topic 1 of topic
2, and so on.

```{r}
df_ex002_list <- extract_student_info(pba_gc_ex002, "preference", 
                                      self_formed_groups = 2, 
                                      pref_mat = pba_prefmat_ex002)
yaml_ex002_list <- extract_params_yaml(system.file("extdata", 
                                             "pba_params_ex002.yml",  
                                             package = "grouper"),
                                       "preference")
m2 <- prepare_model(df_ex002_list, yaml_ex002_list, "preference")

#result2 <- solve_model(m2, with_ROI(solver="gurobi"))
result2 <- solve_model(m2, with_ROI(solver="glpk"))
assign_groups(result2, assignment = "preference", 
              dframe=pba_gc_ex002, yaml_ex002_list, 
              group_names="grouping")
```

# PhD Workload Assignment

In the following walk through, we demonstrate how the PhD work allocation model
can be used on a simple dataset. 

## Dataset 001(Year-long)

This example dataset has 4 students and 4 courses. Each student has non-zero
past workload with `past_ta + past_gr = 4`

```{r}
phd_students_ex001
```

Each row in the preference matrix encodes first/second/third choices as `3/2/1`, 
while all other courses are set to `-99`. Note that different numbering can be
used. 

```{r}
phd_prefmat_ex001
```

The demand table includes only `TA` and `GR`, and intentionally excludes `E`.
Since TA+GR demand is below total student capacity, we can use `e_mode = "rr"`
to automatically generate E demand.

```{r}
phd_demand_ex001
```



```{r}
c_sem <- 4
c(capacity_units = nrow(phd_students_ex001) * c_sem,
  ta_gr_demand = sum(phd_demand_ex001))
```

The following chunk focuses on the three core PhD workflow functions:

1.  `extract_phd_info()`
2.  `prepare_model(..., assignment = "phd")`
3.  `assign_job()`

```{r}
phd_ex001_list <- extract_phd_info(
  student_df = phd_students_ex001,
  p_mat = phd_prefmat_ex001,
  d_mat = phd_demand_ex001,
  e_mode = "rr",
  C = c_sem
)

# show generated demand matrix with E appended by round-robin
phd_ex001_list$d
```

```{r}
m_phd_ex001 <- prepare_model(
  phd_ex001_list,
  assignment = "phd",
  t_max_y1 = 1,
  ta_min = 1, ta_max = 1,
  gr_min = 1, gr_max = 1,
  alpha = 2, beta = 1, phi = 1, rho = 10,
  C = c_sem
)
```


```{r}
result_phd_ex001 <- solve_model(m_phd_ex001, with_ROI(solver = "glpk"))

assign_job(
  result_phd_ex001,
  student_df = phd_students_ex001,
  course_codes = rownames(phd_demand_ex001),
  name_col = "Name"
)
```

## Dataset 001 (Single-semester)

The model can be used to allocate a single semester if we set `past_ta = 0` and 
`past_gr = C` for every student. By doing so, past workload will not contribute 
to the Spread calculation in the objective function. 

We will reuse `phd_students_ex001`.

```{r}
phd_students_ex001_sem <- phd_students_ex001
phd_students_ex001_sem$past_ta <- 0
phd_students_ex001_sem$past_gr <- c_sem

phd_students_ex001_sem
```

```{r}
phd_ex001_sem_list <- extract_phd_info(
  student_df = phd_students_ex001_sem,
  p_mat = phd_prefmat_ex001,
  d_mat = phd_demand_ex001,
  e_mode = "rr",
  C = c_sem
)

phd_ex001_sem_list$d
```

```{r}
m_phd_ex001_sem <- prepare_model(
  phd_ex001_sem_list,
  assignment = "phd",
  t_max_y1 = 1,
  ta_min = 1, ta_max = 1,
  gr_min = 1, gr_max = 1,
  alpha = 2, beta = 1, phi = 1, rho = 10,
  C = c_sem
)

result_phd_ex001_sem <- solve_model(m_phd_ex001_sem, with_ROI(solver = "glpk"))

job_phd_ex001_sem <- assign_job(
  result_phd_ex001_sem,
  student_df = phd_students_ex001_sem,
  course_codes = rownames(phd_demand_ex001),
  name_col = "Name"
)

job_phd_ex001_sem
```

```{r}
# quick summary of each student and worktype
ta_cols <- grepl("-t$", names(job_phd_ex001_sem))
gr_cols <- grepl("-g$", names(job_phd_ex001_sem))
e_cols <- grepl("-e$", names(job_phd_ex001_sem))

data.frame(
  Name = job_phd_ex001_sem$Name,
  TA = rowSums(job_phd_ex001_sem[, ta_cols, drop = FALSE]),
  GR = rowSums(job_phd_ex001_sem[, gr_cols, drop = FALSE]),
  E = rowSums(job_phd_ex001_sem[, e_cols, drop = FALSE]),
  current_total = rowSums(job_phd_ex001_sem[, ta_cols | gr_cols | e_cols, drop = FALSE])
)
```
Note that this is done to showcase the functionality of the model and in this 
simple test data, the result is the same.

The four hyperparameter inside the objective function can be turned off by
setting the value to be 0.

Note: Standalone constructors are also available:
`prepare_diversity_model()`, `prepare_preference_model()`,
and `prepare_phd_model()`.
