---
title: "Related work"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Related work}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
--- 


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(groupr)
```

## Introduction

In this vignette, we provide a brief, non-exhaustive overview of related work to 
assign students to projects, in an educational setting. 

## Balancing staff workload and student preferences

One of the earlier papers found was by @anwar2003student. The goal of the educator
in that paper was to assign students to two types of project: individual ones 
and group ones. For the individual projects, students were asked to rank their top
four projects from a list of project briefs. Each student was then assigned to a 
project and a corresponding staff member.

The assignment was carried out using two models. In the first model, the
objective was to minimise the total number of projects that each staff managed.
In the second model, the feasible solutions from the first model were used to
decide upon an assignment that maximised the number of students obtaining their
first choice projects. 

Compared to the models in `groupr`, the primary difference is in the objective
function. In our case, we only work with the student preference; there is no
need to balance staff workload since the projects are run within a single
course.

For the group project, students were asked to rank their top 3 projects
*individually*. They were then allocated groups that maximised their preferences.
Compared to the model in our package, this approach does not allow for
self-formed groups, or for larger projects that involve sub-groups.

## OptAssign

In many ways, the model described in @meyer2009optassign is similar to ours. The
objective function there maximises student preference for project topics. The
model in the paper also discusses an extension whereby some groups of students
can be favoured for particular topics. This is carried out through the inclusion
of weights in the objective function. Another additional feature adds
contstraints to ensure that students with certain skills are included in
particular topics. For instance, a project topic on "Marketing surveys" may
require at least two students with a statistics background in the group.

Overall, this model formulation is very similar to the one in our package.
Again, one difference is in the affordance for self-formed groups, and for
sub-groups in the project team.

## Students to elective courses

Let us try a reference here @meyer2009optassign. Another reference @anwar2003student,

## References

