---
title: "PhD model testing AY 2520"
format: html
---

```{r}
library(tidyverse)
library(ompr)
library(ompr.roi)
library(ROI.plugin.gurobi)
library(ROI.plugin.glpk)
library(ROI.plugin.highs)
library(devtools)
library(readxl)
library(digest)
library(purrr)
```

```{r}
# Setting up
load_all()
exists("prepare_phd_model", where = globalenv(), inherits = FALSE)
source("phdmodel_test_helpers.R")
```


## Loading Data

```{r}
preference <- read_excel("../data/phdModel/AY2520/preference.xlsx", range = "B4:O44")
demand <- read_excel("../data/phdModel/AY2520/demand.xlsx")
past_output_2510 <- read_excel("../data/phdModel/AY2510/output_005.xlsx",sheet = "Sheet3")
```

## Data cleaning

```{r}
# parameters
Ns <- nrow(preference)
Nj <- nrow(demand)
C <- 4
```

```{r}
# Date handling
reference_date <- ymd("2026-05-01")

preference <- preference %>%
  mutate(intake_month = case_when(
    str_detect(`Year (intake enrolled in)`, "Jan") ~ 1,
    str_detect(`Year (intake enrolled in)`, "Aug") ~ 8,
    TRUE ~ NA),
    intake_year = as.integer(str_extract(`Year (intake enrolled in)`, "\\d{4}")),
    intake_date = ymd(paste(intake_year, intake_month, "01")),
    year_raw = time_length(interval(intake_date, reference_date), "years"),
    seniority = ceiling(year_raw),
    Name = standardise_name(Name)
  ) %>%
  rename(first = 9, second= 10, third = 11) %>%
  arrange(`S/No.`)

# Student-id mapping
student_id <- preference %>%
  select(`S/No.`, Name, seniority) %>%
  rename(ID = `S/No.`)
```

```{r}
# Extracting demand
demand_old <- demand # save a old copy for debugging

demand <- demand %>%
  rename(TA = ta_hours, GR = grading_hours) %>%
  mutate(E = numeric(Nj)) %>%
  arrange(ID)

################# HARD CODING CHANGE IN TA HOUR ##############
demand[12, 5] = demand[12, 5] - 8 # ST2334
demand[8,5] = demand[8,5] - 5 # ST1131
############################################################## 

# calculate the number of Empty units
total_units <- Ns * C
total_demand <- sum(demand[["TA"]]) + sum(demand[["GR"]])
total_E <- total_units - total_demand

# Allocate E proportionally(RR fashion)
eligible <- which(demand$GR > 0)
ranked <- eligible[order(demand$GR[eligible], decreasing = TRUE)]
rr_idx <- ranked[((seq_len(total_E) - 1) %% length(ranked)) + 1]
demand$E <- tabulate(rr_idx, nbins = nrow(demand))

# converting to matrix
D <- as.matrix(demand[,c("TA", "GR", "E")])

# building look up course list
courses <- norm_code(pull(demand, course_code))
```
**note that it is very important to ensure P[,j] and D[j,] are referring to the 
same course.**

```{r}
# preference matrix
P <- matrix(-99L, nrow = Ns, ncol = Nj)

for (i in seq_len(Ns)) {

  first  <- norm_code(preference[["first"]][i])
  second <- norm_code(preference[["second"]][i])
  third  <- norm_code(preference[["third"]][i])

  # if first choice is NA, all courses = 3
  if (is.na(first) || first == "") {
    P[i, ] <- 3L
    next
  }

  if (!is.na(first) && first != "") {
    j <- match(first, courses)
    if (!is.na(j)) P[i, j] <- 3L
  }

  if (!is.na(second) && second != "") {
    j <- match(second, courses)
    if (!is.na(j)) P[i, j] <- 2L
  }

  if (!is.na(third) && third != "") {
    j <- match(third, courses)
    if (!is.na(j)) P[i, j] <- 1L
  }
}
```


```{r}
# Extracting previous semester workload
past_ta_cols <- names(past_output_2510)[str_detect(names(past_output_2510), "-t$")]
past_gr_cols <- names(past_output_2510)[str_detect(names(past_output_2510), "-g$")]

# Past workload for AY2520 is sourced from AY2510 output_005.xlsx.
past_summary <- past_output_2510 %>%
  filter(!is.na(Name), Name != "") %>%
  mutate(
    Name = standardise_name(Name),
    TA = rowSums(across(all_of(past_ta_cols)), na.rm = TRUE),
    GR = rowSums(across(all_of(past_gr_cols)), na.rm = TRUE)
  ) %>%
  group_by(Name) %>%
  summarise(
    TA = sum(TA),
    GR = sum(GR),
    .groups = "drop"
  )

# AY2520 students with no AY2510 rows default to zero before final padding.
past_clean <- student_id %>%
  left_join(past_summary, by = "Name") %>%
  mutate(
    TA = replace_na(TA, 0),
    GR = replace_na(GR, 0)
  ) %>%
  arrange(ID)
  
  
# check if every one has exactly C units in total. If less than C, pad with GR.
past_final <- past_clean %>%
  mutate(
    total = TA + GR,
    pad_GR = pmax(0, C - total),
    GR = GR + pad_GR
  ) %>%
  select(ID, Name, TA, GR) %>%
  arrange(ID)
```
Note: during actual package implementation, ensure every student exists in past 
table (even if they had no past assignments). remove those not in current 
allocation.

```{r}
# Input prep via package extractor
student_input <- student_id %>%
  left_join(past_final %>% select(ID, TA, GR), by = "ID") %>%
  transmute(
    student_id = ID,
    year = seniority,
    past_ta = TA,
    past_gr = GR
  )

df_list <- extract_phd_info(
  student_df = student_input,
  p_mat = P,
  d_mat = D,
  e_mode = "none",
  C = C
)

s <- df_list$s
t1 <- df_list$t1
g1 <- df_list$g1
```

## Modelling
```{r}
# initialisation
model_0 <- prepare_phd_model(df_list, e_max = 1, C = C)

result_0 <- solve_model(model_0, with_ROI(solver="gurobi"))

alloc_df_0 <- extract_phd_assignment(model_result = result_0, 
                                     student_df = preference)

plot_phd_year_distribution(alloc_df_0, 
                           t1 = df_list$t1, g1 = df_list$g1,
                           facet_by_seniority = TRUE, seniority_col = "seniority")
```


```{r}
get_solution(result_0, Tmax);get_solution(result_0, Tmin) 

# can try change alpha from default to 100. Better spread but slow down the solver. 
```

## Retrieving top-k result
https://docs.gurobi.com/projects/optimizer/en/current/features/solutionpool.html
```{r}
# build solution pool
result_pool <- solve_model(
  model_0,
  with_ROI(
    solver = "gurobi",
    PoolSearchMode = 2,
    PoolSolutions = 50,
    verbose = TRUE
  )
)
```

```{r}
pool_results <- extract_phd_pool_results(
  result_pool = result_pool,
  preference = preference,
  k = 50
)
```

```{r}
pool_results[[2]]$phd_assignment
```

```{r}
pool_results[[2]]$ta_detail
```

```{r}
# top 3 unique
n_plot <- 3

for (i in seq_len(n_plot)) {
  cur <- pool_results[[i]]

  p <- plot_phd_year_distribution(
    current_alloc_df = cur$phd_assignment,
    t1 = df_list$t1,
    g1 = df_list$g1,
    facet_by_seniority = TRUE,
    seniority_col = "seniority",
    facet_ncol = 4
  ) +
    ggplot2::labs(
      subtitle = paste0(
        "rank=", i,
        " | obj=", round(cur$objective_value, 4),
        " | TA min=", round(cur$ta_min, 4),
        " | TA max=", round(cur$ta_max, 4)
      )
    )

  print(p)
}
```


## Using other solver 

```{r}
bench_solver <- function(model, solver_name) {
  t <- system.time({
    res <- solve_model(model, with_ROI(solver = solver_name))
  })
  list(
    solver = solver_name,
    result = res,
    time_sec = unname(t["elapsed"])
  )
}

b_gurobi <- bench_solver(model_0, "gurobi")
b_glpk   <- bench_solver(model_0, "glpk")
b_highs  <- bench_solver(model_0, "highs")

solver_compare <- tibble::tibble(
  solver = c("gurobi", "glpk", "highs"),
  time_sec = c(b_gurobi$time_sec, b_glpk$time_sec, b_highs$time_sec),
  objective = c(
    objective_value(b_gurobi$result),
    objective_value(b_glpk$result),
    objective_value(b_highs$result)
  )
)

solver_compare
```

```{r}
alloc_df_glpk <- extract_phd_assignment(
  model_result = b_glpk[[2]],
  student_df = preference
)

alloc_df_highs <- extract_phd_assignment(
  model_result = b_highs[[2]],
  student_df = preference
)


plot_phd_year_distribution(alloc_df_glpk,t1 = df_list$t1, g1 = df_list$g1,
  facet_by_seniority = TRUE, seniority_col = "seniority")

plot_phd_year_distribution(alloc_df_highs,
  t1 = df_list$t1, g1 = df_list$g1,
  facet_by_seniority = TRUE, seniority_col = "seniority"
)

plot_phd_year_distribution(alloc_df_0, t1 = df_list$t1, g1 = df_list$g1,
                           facet_by_seniority = TRUE, seniority_col = "seniority")
```


## Objective Comparison with manual output

```{r}
manual_output <- read_excel("../data/phdModel/AY2520/output_001.xlsx",sheet = "worker_tbl_2520") %>%
  mutate(name_std = standardise_name(Name))

pref_lookup <- preference %>%
  transmute(
    i = row_number(),
    student_id = `S/No.`,
    student_name = Name,
    name_std = standardise_name(Name)
  )

manual_alloc <- pref_lookup %>%
  left_join(
    manual_output %>%
      select(name_std, tutoring, grading),
    by = "name_std"
  ) %>%
  transmute(
    student_id,
    student_name,
    manual_TA = replace_na(as.numeric(tutoring), 0),
    manual_GR = replace_na(as.numeric(grading), 0),
    manual_E = pmax(0, 4 - manual_TA - manual_GR)
  )
```

```{r}
# Must match the settings used in prepare_phd_model above.
obj_cfg <- list(alpha = 2, beta = 1, phi = 1, rho = 10, t_max_y1 = 1)
```

```{r}
# Solver objective for the model solution.
model_solver_obj <- as.numeric(result_0$objective_value)
model_solver_obj <- model_solver_obj[[1]]
```

```{r}
# Recalculate manual objective using manual TA by course + derived E.
manual_ta_detail <- manual_output %>%
  inner_join(pref_lookup, by = "name_std") %>%
  select(i, student_id, student_name, matches("-t$")) %>%
  pivot_longer(
    cols = matches("-t$"),
    names_to = "course_col",
    values_to = "ta_units"
  ) %>%
  mutate(
    ta_units = replace_na(as.numeric(ta_units), 0),
    course_code = norm_code(str_remove(course_col, "-t$")),
    j = match(course_code, courses)
  ) %>%
  filter(ta_units > 1e-8)

manual_ta_pref_sum <- manual_ta_detail %>%
  filter(!is.na(j)) %>%
  mutate(pref_score = P[cbind(i, j)]) %>%
  summarise(val = sum(pref_score * ta_units, na.rm = TRUE)) %>%
  pull(val)

manual_totals_by_i <- pref_lookup %>%
  select(i, student_id, student_name) %>%
  left_join(manual_alloc, by = c("student_id", "student_name")) %>%
  mutate(
    manual_TA = replace_na(as.numeric(manual_TA), 0),
    manual_E = replace_na(as.numeric(manual_E), 0)
  ) %>%
  arrange(i)
```

```{r}
manual_obj_breakdown <- compute_obj_from_components(
  ta_by_student = manual_totals_by_i$manual_TA,
  e_by_student = manual_totals_by_i$manual_E,
  ta_pref_sum = manual_ta_pref_sum,
  t1 = t1,
  s = s,
  t_max_y1 = obj_cfg$t_max_y1,
  alpha = obj_cfg$alpha,
  beta = obj_cfg$beta,
  phi = obj_cfg$phi,
  rho = obj_cfg$rho
)

manual_obj_breakdown
```

```{r}
model_solver_obj
```

## Using package function to get Solver Output

```{r}
job_output <- assign_job(
  model_result = result_0,
  student_df = preference,
  course_codes = courses,
  name_col = "Name"
)

job_output
```

```{r}
# check per-student TA / GR sessions from job_output
ta_cols_job <- names(job_output)[str_detect(names(job_output), "-t$")]
gr_cols_job <- names(job_output)[str_detect(names(job_output), "-g$")]
e_cols_job <- names(job_output)[str_detect(names(job_output), "-e$")]

job_output_summary <- job_output %>%
  transmute(
    Name,
    TA_sessions = rowSums(across(all_of(ta_cols_job)), na.rm = TRUE),
    GR_sessions = rowSums(across(all_of(gr_cols_job)), na.rm = TRUE),
    E_sessions = rowSums(across(all_of(e_cols_job)), na.rm = TRUE)
  ) %>%
  arrange(Name)

job_output_summary
```
