---
title: "PhD model testing AY 2510"
format: html
---

```{r}
library(tidyverse)
library(ompr)
library(ompr.roi)
library(ROI.plugin.gurobi)
library(devtools)
library(readxl)
library(lubridate)
library(digest)
library(purrr)
```

```{r}
# Setting up
load_all()
exists("prepare_phd_model", where = globalenv(), inherits = FALSE)
source("phdmodel_test_helpers.R")
```

## Loading Data

```{r}
preference_y24_raw <- read_excel(
  "../data/phdModel/AY2510/List of research scholars and teaching preferences for Sem 1, AY202526.xlsx",
  sheet = "List of RS ( Yr 2-4 )", skip = 2
)

preference_y1_raw <- read_excel(
  "../data/phdModel/AY2510/List of research scholars and teaching preferences for Sem 1, AY202526.xlsx",
  sheet = "List of RS ( Yr 1 )", skip = 2
)

manual_output_current <- read_excel("../data/phdModel/AY2510/output_005.xlsx", sheet = "Sheet3")
past_output_2420 <- read_excel("../data/phdModel/AY2420/output_009.xlsx", sheet = "output_001a")
```

## Data cleaning

```{r}
# Date handling
reference_date <- ymd("2025-12-01")

preference_y24 <- preference_y24_raw %>%
  filter(!is.na(`S/No.`)) %>%
  transmute(
    `S/No._raw` = as.integer(`S/No.`),
    Name,
    `Year (intake enrolled in)` = as.character(`Year (intake enrolled in)`),
    first = clean_pref_code(`Please provide the module code which is your first tutorial preferences (with 1 being most preferred) (first choice)`),
    second = clean_pref_code(`Please list the module code which is your second tutorial preferences (second choice)`),
    third = clean_pref_code(`Please list the module code which is your third tutorial preferences (Third choice)`),
    intake_month = case_when(
      str_detect(str_to_lower(`Year (intake enrolled in)`), "jan") ~ 1L,
      str_detect(str_to_lower(`Year (intake enrolled in)`), "aug") ~ 8L,
      TRUE ~ NA_integer_
    ),
    intake_year = as.integer(str_extract(`Year (intake enrolled in)`, "\\d{4}")),
    intake_date = ymd(paste(intake_year, intake_month, "01")),
    source_sheet = "Yr2-4"
  )

preference_y1 <- preference_y1_raw %>%
  filter(!is.na(`S/No.`)) %>%
  mutate(intake_date = as.Date(`Year (intake enrolled in)`)) %>%
  transmute(
    `S/No._raw` = as.integer(`S/No.`),
    Name,
    `Year (intake enrolled in)` = as.character(intake_date),
    first = NA_character_,
    second = NA_character_,
    third = NA_character_,
    intake_month = month(intake_date),
    intake_year = year(intake_date),
    intake_date,
    source_sheet = "Yr1"
  )
```

```{r}
# AY2510 has overlapping S/No. across Yr1 and Yr2-4 sheets.
# Rebuild a unique student ID after combining both sheets.
preference <- bind_rows(preference_y24, preference_y1) %>%
  mutate(
    year_raw = time_length(interval(intake_date, reference_date), "years"),
    seniority = ceiling(year_raw),
    Name = standardise_name(Name)
  ) %>%
  # cap seniority at 4 (treat year 5/6 as year 4)
  mutate(seniority = pmin(seniority, 4L)) %>%
  arrange(source_sheet, `S/No._raw`, Name) %>%
  mutate(`S/No.` = row_number()) %>%
  relocate(`S/No.`, .before = Name)


# Student-id mapping
student_id <- preference %>%
  select(`S/No.`, Name, seniority) %>%
  rename(ID = `S/No.`)
```

```{r}
# Extracting demand
demand_source <- manual_output_current %>%
  filter(str_detect(email, "@"), !is.na(Name)) %>%
  mutate(name_std = standardise_name(Name))

demand_ta_cols <- names(demand_source)[str_detect(names(demand_source), "-t$")]
demand_gr_cols <- names(demand_source)[str_detect(names(demand_source), "-g$")]

# Demand is now derived directly from manually completed output_005.xlsx.
# Exclude full-time TA WONG YEAN LING from the remaining allocatable demand.
demand_allocatable <- demand_source %>%
  filter(name_std != "WONG YEAN LING")

demand_ta <- demand_allocatable %>%
  summarise(across(all_of(demand_ta_cols), ~ sum(as.numeric(.x), na.rm = TRUE))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "course_col",
    values_to = "TA"
  ) %>%
  transmute(
    course_code = norm_code(str_remove(course_col, "-t$")),
    TA = replace_na(as.numeric(TA), 0)
  )

demand_gr <- demand_allocatable %>%
  summarise(across(all_of(demand_gr_cols), ~ sum(as.numeric(.x), na.rm = TRUE))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "course_col",
    values_to = "GR"
  ) %>%
  transmute(
    course_code = norm_code(str_remove(course_col, "-g$")),
    GR = replace_na(as.numeric(GR), 0)
  )

demand <- full_join(demand_ta, demand_gr, by = "course_code") %>%
  mutate(
    TA = replace_na(TA, 0),
    GR = replace_na(GR, 0),
    E = 0 # E not derived from output_005 in this workflow
  ) %>%
  filter(TA + GR > 0) %>%
  arrange(course_code) %>%
  mutate(ID = row_number()) %>%
  relocate(ID, .before = course_code)
```

```{r}
# parameters
Ns <- nrow(preference)
Nj <- nrow(demand)
C <- 4
```

```{r}
# calculate the number of Empty units
total_units <- Ns * C
total_demand <- sum(demand[["TA"]]) + sum(demand[["GR"]])
total_E <- total_units - total_demand

# Allocate E in round robin fashion from highest to lowest GR demand.
eligible <- which(demand$GR > 0)

if (total_E > 0 && length(eligible) > 0) {
  ranked <- eligible[order(demand$GR[eligible], decreasing = TRUE)]
  rr_idx <- ranked[((seq_len(total_E) - 1) %% length(ranked)) + 1]
  demand$E <- tabulate(rr_idx, nbins = nrow(demand))
}

# converting to matrix
D <- as.matrix(demand[, c("TA", "GR", "E")])

# building look up course list
courses <- norm_code(pull(demand, course_code))
```
**note that it is very important to ensure P[,j] and D[j,] are referring to the
same course.**

```{r}
# preference matrix
P <- matrix(-99L, nrow = Ns, ncol = Nj)

for (i in seq_len(Ns)) {
  first <- norm_code(preference[["first"]][i])
  second <- norm_code(preference[["second"]][i])
  third <- norm_code(preference[["third"]][i])

  # if first choice is NA, all courses = 3
  if (is.na(first) || first == "") {
    P[i, ] <- 3L
    next
  }

  if (!is.na(first) && first != "") {
    j <- match(first, courses)
    if (!is.na(j)) P[i, j] <- 3L
  }

  if (!is.na(second) && second != "") {
    j <- match(second, courses)
    if (!is.na(j)) P[i, j] <- 2L
  }

  if (!is.na(third) && third != "") {
    j <- match(third, courses)
    if (!is.na(j)) P[i, j] <- 1L
  }
}
```

```{r}
# Extracting previous semester workload
past_ta_cols <- names(past_output_2420)[str_detect(names(past_output_2420), "-t$")]
past_gr_cols <- names(past_output_2420)[str_detect(names(past_output_2420), "-g$")]

# Past workload is now derived from AY2420 output_009.xlsx.
past_summary <- past_output_2420 %>%
  filter(!is.na(Name), Name != "") %>%
  mutate(
    Name = standardise_name(Name),
    TA = rowSums(across(all_of(past_ta_cols)), na.rm = TRUE),
    GR = rowSums(across(all_of(past_gr_cols)), na.rm = TRUE)
  ) %>%
  group_by(Name) %>%
  summarise(
    TA = sum(TA),
    GR = sum(GR),
    .groups = "drop"
  )
```

```{r}
# AY2420-only students are not included in this workflow.
# AY2510 students with no past rows default to zero before final padding.
past_clean <- student_id %>%
  left_join(past_summary, by = "Name") %>%
  mutate(
    TA = replace_na(TA, 0),
    GR = replace_na(GR, 0)
  ) %>%
  arrange(ID)


# check if every one has exactly C units in total. If less than C, pad with GR.
past_final <- past_clean %>%
  mutate(
    total = TA + GR,
    pad_GR = pmax(0, C - total),
    GR = GR + pad_GR
  ) %>%
  select(ID, Name, TA, GR) %>%
  arrange(ID)
```

Note: during actual package implementation, ensure every student exists in past
table (even if they had no past assignments). remove those not in current
allocation.

```{r}
# Extracting seniority
s <- pull(preference, seniority) - 2 # offset
```

```{r}
# Input prep
t1 <- pull(past_final, TA)
g1 <- pull(past_final, GR)

# final input to model
df_list = list(
  Ns = Ns,
  Nj = Nj,
  P = P,
  d = D,
  s = s,
  t1 = t1,
  g1 = g1
)
```

## Modelling
```{r}
# initialisation
model_0 <- prepare_phd_model(df_list, e_max = 2, C = C)

result_0 <- solve_model(model_0, with_ROI(solver = "gurobi"))

alloc_df_0 <- extract_phd_assignment(
  model_result = result_0,
  student_df = preference
)

plot_phd_year_distribution(
  alloc_df_0,
  t1 = df_list$t1, g1 = df_list$g1,
  facet_by_seniority = TRUE, seniority_col = "seniority"
)
```

```{r}
get_solution(result_0, Tmax); get_solution(result_0, Tmin)
```

```{r}
# Preference fulfilment
ta_detail <- extract_phd_ta_detail(result_0, preference)
```

## Retrieving top-k result
https://docs.gurobi.com/projects/optimizer/en/current/features/solutionpool.html
```{r}
# build solution pool
result_pool <- solve_model(
  model_0,
  with_ROI(
    solver = "gurobi",
    PoolSearchMode = 2,
    PoolSolutions = 50,
    verbose = TRUE
  )
)
```
```{r}
pool_results <- extract_phd_pool_results(
  result_pool = result_pool,
  preference = preference,
  k = 50
)
```

```{r}
pool_results[[2]]$phd_assignment
```

```{r}
pool_results[[2]]$ta_detail
```

```{r}
# top 3 unique
n_plot <- 3

for (i in seq_len(n_plot)) {
  cur <- pool_results[[i]]

  p <- plot_phd_year_distribution(
    current_alloc_df = cur$phd_assignment,
    t1 = df_list$t1,
    g1 = df_list$g1,
    facet_by_seniority = TRUE,
    seniority_col = "seniority",
    facet_ncol = 4
  ) +
    ggplot2::labs(
      subtitle = paste0(
        "rank=", i,
        " | obj=", round(cur$objective_value, 4),
        " | TA min=", round(cur$ta_min, 4),
        " | TA max=", round(cur$ta_max, 4)
      )
    )

  print(p)
}
```

## Objective Comparison with manual output

```{r}
manual_output <- demand_source

pref_lookup <- preference %>%
  transmute(
    i = row_number(),
    student_id = `S/No.`,
    student_name = Name,
    name_std = standardise_name(Name)
  )

ta_cols <- demand_ta_cols
gr_cols <- demand_gr_cols

manual_alloc <- pref_lookup %>%
  left_join(
    manual_output %>%
      transmute(
        name_std,
        manual_TA = rowSums(across(all_of(ta_cols)), na.rm = TRUE),
        manual_GR = rowSums(across(all_of(gr_cols)), na.rm = TRUE)
      ) %>%
      distinct(name_std, .keep_all = TRUE),
    by = "name_std"
  ) %>%
  transmute(
    student_id,
    student_name,
    manual_TA = replace_na(as.numeric(manual_TA), 0),
    manual_GR = replace_na(as.numeric(manual_GR), 0),
    manual_E = pmax(0, 4 - manual_TA - manual_GR)
  )
```

```{r}
# Must match the settings used in prepare_phd_model above.
obj_cfg <- list(alpha = 2, beta = 1, phi = 1, rho = 10, t_max_y1 = 1)
```

```{r}
# Solver objective for the model solution.
model_solver_obj <- as.numeric(result_0$objective_value)
if (length(model_solver_obj) == 0 || is.na(model_solver_obj[[1]])) {
  model_solver_obj <- as.numeric(result_0$additional_solver_output$ROI$message$objval)
}
model_solver_obj <- model_solver_obj[[1]]
```

```{r}
# Recalculate manual objective using manual TA by course + derived E.
manual_ta_detail <- manual_output %>%
  inner_join(pref_lookup, by = "name_std") %>%
  select(i, student_id, student_name, matches("-t$")) %>%
  pivot_longer(
    cols = matches("-t$"),
    names_to = "course_col",
    values_to = "ta_units"
  ) %>%
  mutate(
    ta_units = replace_na(as.numeric(ta_units), 0),
    course_code = norm_code(str_remove(course_col, "-t$")),
    j = match(course_code, courses)
  ) %>%
  filter(ta_units > 1e-8)

manual_ta_pref_sum <- manual_ta_detail %>%
  filter(!is.na(j)) %>%
  mutate(pref_score = P[cbind(i, j)]) %>%
  summarise(val = sum(pref_score * ta_units, na.rm = TRUE)) %>%
  pull(val)

manual_totals_by_i <- pref_lookup %>%
  select(i, student_id, student_name) %>%
  left_join(manual_alloc, by = c("student_id", "student_name")) %>%
  mutate(
    manual_TA = replace_na(as.numeric(manual_TA), 0),
    manual_E = replace_na(as.numeric(manual_E), 0)
  ) %>%
  arrange(i)
```

```{r}
manual_obj_breakdown <- compute_obj_from_components(
  ta_by_student = manual_totals_by_i$manual_TA,
  e_by_student = manual_totals_by_i$manual_E,
  ta_pref_sum = manual_ta_pref_sum,
  t1 = t1,
  s = s,
  t_max_y1 = obj_cfg$t_max_y1,
  alpha = obj_cfg$alpha,
  beta = obj_cfg$beta,
  phi = obj_cfg$phi,
  rho = obj_cfg$rho
)

manual_obj_breakdown
```

```{r}
model_solver_obj
```

## Manual-Style Solver Output

```{r}
job_output <- assign_job(
  model_result = result_0,
  student_df = preference,
  course_codes = courses,
  name_col = "Name"
)

job_output
```

```{r}
# check per-student TA / GR sessions from job_output
ta_cols_job <- names(job_output)[str_detect(names(job_output), "-t$")]
gr_cols_job <- names(job_output)[str_detect(names(job_output), "-g$")]
e_cols_job <- names(job_output)[str_detect(names(job_output), "-e$")]

job_output_summary <- job_output %>%
  transmute(
    Name,
    TA_sessions = rowSums(across(all_of(ta_cols_job)), na.rm = TRUE),
    GR_sessions = rowSums(across(all_of(gr_cols_job)), na.rm = TRUE),
    E_sessions = rowSums(across(all_of(e_cols_job)), na.rm = TRUE)
  ) %>%
  arrange(Name)

job_output_summary
```
