---
title: "PhD model testing AY 2420"
format: html
---

```{r}
library(tidyverse)
library(ompr)
library(ompr.roi)
library(ROI.plugin.gurobi)
library(devtools)
library(readxl)
library(lubridate)
library(digest)
library(purrr)
```

```{r}
# Setting up
load_all()
exists("prepare_phd_model", where = globalenv(), inherits = FALSE)
source("phdmodel_test_helpers.R")
```

## Loading Data

```{r}
preference_y24_raw <- read_excel(
  "../data/phdModel/AY2420/List of research scholars and teaching preferences for Sem 2 AY2425.xlsx",
  sheet = "List of RS ( Yr 2-4 )",
  skip = 2, range = "B4:O37"
)

preference_y1_raw <- read_excel(
  "../data/phdModel/AY2420/List of research scholars and teaching preferences for Sem 2 AY2425.xlsx",
  sheet = "List of RS ( Yr 1 )",
  skip = 2
)

demand_raw <- read_excel("../data/phdModel/AY2420/2425_Sem2_tutor_grader_request_working.xlsx", sheet = 1)
manual_output_current <- read_excel("../data/phdModel/AY2420/output_009.xlsx", sheet = "output_001a")
past_output_2410 <- read_excel("../data/phdModel/AY2410/output_009.xlsx", sheet = "full_tbl")
```


## Data cleaning
```{r}
# Date handling
reference_date <- ymd("2025-05-01")

preference_y24 <- preference_y24_raw %>%
  filter(!is.na(`S/No.`)) %>%
  transmute(
    `S/No._raw` = as.integer(`S/No.`),
    Name,
    `Year (intake enrolled in)` = as.character(`Year (intake enrolled in)`),
    first = clean_pref_code(`Please provide the module code which is your first tutorial preferences (with 1 being most preferred) (first choice)`),
    second = clean_pref_code(`Please list the module code which is your second tutorial preferences (second choice)`),
    third = clean_pref_code(`Please list the module code which is your third tutorial preferences (Third choice)`),
    intake_month = case_when(
      str_detect(str_to_lower(`Year (intake enrolled in)`), "jan") ~ 1L,
      str_detect(str_to_lower(`Year (intake enrolled in)`), "aug") ~ 8L,
      TRUE ~ NA_integer_
    ),
    intake_year = as.integer(str_extract(`Year (intake enrolled in)`, "\\d{4}")),
    intake_date = ymd(paste(intake_year, intake_month, "01")),
    source_sheet = "Yr2-4"
  )

preference_y1 <- preference_y1_raw %>%
  filter(!is.na(`S/No.`)) %>%
  mutate(
    intake_date = as.Date(`Year (intake enrolled in)`),
    # AY2420 Yr1 sheet has missing intake dates it should be Aug-2024.
    intake_date = coalesce(intake_date, ymd("2024-08-01"))
  ) %>%
  transmute(
    `S/No._raw` = as.integer(`S/No.`),
    Name,
    `Year (intake enrolled in)` = as.character(intake_date),
    first = NA_character_,
    second = NA_character_,
    third = NA_character_,
    intake_month = month(intake_date),
    intake_year = year(intake_date),
    intake_date,
    source_sheet = "Yr1"
  )
```

```{r}
# AY2420 has overlapping S/No. across Yr1 and Yr2-4 sheets.
# Rebuild a unique student ID after combining both sheets.
preference <- bind_rows(preference_y24, preference_y1) %>%
  mutate(
    year_raw = time_length(interval(intake_date, reference_date), "years"),
    seniority = ceiling(year_raw),
    Name = standardise_name(Name)
  ) %>%
  # cap seniority at 4 (treat year 5/6 as year 4)
  mutate(seniority = pmin(seniority, 4L)) %>%
  arrange(source_sheet, `S/No._raw`, Name) %>%
  mutate(`S/No.` = row_number()) %>%
  relocate(`S/No.`, .before = Name)


# Student-id mapping
student_id <- preference %>%
  select(`S/No.`, Name, seniority) %>%
  rename(ID = `S/No.`)
```

```{r}
# Extracting demand
demand_cols <- tail(names(demand_raw), 2)

# AY2420 demand source: first sheet's last 2 columns of raw request file.
# GR is reported in hours in this file; convert to model units (11 hrs = 1 unit).
demand <- demand_raw %>%
  transmute(
    course_code = norm_code(`Course Code`),
    TA = suppressWarnings(as.numeric(.data[[demand_cols[1]]])),
    GR = suppressWarnings(as.numeric(.data[[demand_cols[2]]])) / 11
  ) %>%
  mutate(
    TA = replace_na(TA, 0),
    GR = replace_na(GR, 0),
    E = 0
  ) %>%
  filter(!is.na(course_code), course_code != "") %>%
  filter(TA + GR > 0) %>%
  arrange(course_code) %>%
  mutate(ID = row_number()) %>%
  relocate(ID, .before = course_code)

################# HARD CODING CHANGE IN TA HOUR ##############
demand[7, 3] = demand[7, 3] - 5 # ST1131
demand[10,3] = demand[10,3] - 7 # ST2334
demand[5,3] = demand[5,3] - 11 # DSA4265
############################################################## 

```

```{r}
# parameters
Ns <- nrow(preference)
Nj <- nrow(demand)
C <- 4
```


```{r}
# calculate the number of Empty units
total_units <- Ns * C
total_demand <- sum(demand[["TA"]]) + sum(demand[["GR"]])
total_E <- total_units - total_demand

# Allocate E in round robin fashion from highest to lowest GR demand.
eligible <- which(demand$GR > 0)

if (total_E > 0 && length(eligible) > 0) {
  ranked <- eligible[order(demand$GR[eligible], decreasing = TRUE)]
  rr_idx <- ranked[((seq_len(total_E) - 1) %% length(ranked)) + 1]
  demand$E <- tabulate(rr_idx, nbins = nrow(demand))
}

# converting to matrix
D <- as.matrix(demand[, c("TA", "GR", "E")])

# building look up course list
courses <- norm_code(pull(demand, course_code))
```
**note that it is very important to ensure P[,j] and D[j,] are referring to the
same course.**

```{r}
# preference matrix
P <- matrix(-99L, nrow = Ns, ncol = Nj)

# AY2420 demand has cross-listed codes (e.g. ST2131/MA2116).
match_course <- function(code, courses) {
  code <- norm_code(code)
  j <- match(code, courses)
  if (!is.na(j)) return(j)
  starts <- which(str_starts(courses, paste0(code, "/")))
  if (length(starts) == 1) return(starts[[1]])
  NA_integer_
}

for (i in seq_len(Ns)) {
  first <- norm_code(preference[["first"]][i])
  second <- norm_code(preference[["second"]][i])
  third <- norm_code(preference[["third"]][i])

  # if first choice is NA, all courses = 3
  if (is.na(first) || first == "") {
    P[i, ] <- 3L
    next
  }

  if (!is.na(first) && first != "") {
    j <- match_course(first, courses)
    if (!is.na(j)) P[i, j] <- 3L
  }

  if (!is.na(second) && second != "") {
    j <- match_course(second, courses)
    if (!is.na(j)) P[i, j] <- 2L
  }

  if (!is.na(third) && third != "") {
    j <- match_course(third, courses)
    if (!is.na(j)) P[i, j] <- 1L
  }
}
```

```{r}
# Extracting previous semester workload
past_ta_cols <- names(past_output_2410)[str_detect(names(past_output_2410), "-t$")]
past_gr_cols <- names(past_output_2410)[str_detect(names(past_output_2410), "-g$")]

# Past workload for AY2420 is sourced from prior-cycle AY2410 output.
past_summary <- past_output_2410 %>%
  filter(!is.na(Name), Name != "") %>%
  mutate(
    Name = standardise_name(Name),
    TA = rowSums(across(all_of(past_ta_cols)), na.rm = TRUE),
    GR = rowSums(across(all_of(past_gr_cols)), na.rm = TRUE)
  ) %>%
  group_by(Name) %>%
  summarise(
    TA = sum(TA),
    GR = sum(GR),
    .groups = "drop"
  )
```

```{r}
# AY2420 students with no AY2410 rows default to zero before final padding.
past_clean <- student_id %>%
  left_join(past_summary, by = "Name") %>%
  mutate(
    TA = replace_na(TA, 0),
    GR = replace_na(GR, 0)
  ) %>%
  arrange(ID)


# check if every one has exactly C units in total. If less than C, pad with GR.
past_final <- past_clean %>%
  mutate(
    total = TA + GR,
    pad_GR = pmax(0, C - total),
    GR = GR + pad_GR
  ) %>%
  select(ID, Name, TA, GR) %>%
  arrange(ID)
```

Note: during actual package implementation, ensure every student exists in past
table (even if they had no past assignments). remove those not in current
allocation.

```{r}
# Input prep
student_input <- student_id %>%
  left_join(past_final %>% select(ID, TA, GR), by = "ID") %>%
  transmute(
    student_id = ID,
    year = seniority,
    past_ta = TA,
    past_gr = GR
  )

df_list <- extract_phd_info(
  student_df = student_input,
  p_mat = P,
  d_mat = D,
  e_mode = "none",
  C = C
)

s <- df_list$s
t1 <- df_list$t1
g1 <- df_list$g1
```

## Modelling
```{r}
# initialisation
model_0 <- prepare_phd_model(df_list, e_max = 1, C = C)

result_0 <- solve_model(model_0, with_ROI(solver = "gurobi"))

alloc_df_0 <- extract_phd_assignment(
  model_result = result_0,
  student_df = preference
)

plot_phd_year_distribution(
  alloc_df_0,
  t1 = df_list$t1, g1 = df_list$g1,
  facet_by_seniority = TRUE, seniority_col = "seniority"
)
```

```{r}
get_solution(result_0, Tmax); get_solution(result_0, Tmin)
```

```{r}
# Preference fulfilment
ta_detail <- extract_phd_ta_detail(result_0, preference)
```

## Retrieving top-k result
https://docs.gurobi.com/projects/optimizer/en/current/features/solutionpool.html
```{r}
# build solution pool
result_pool <- solve_model(
  model_0,
  with_ROI(
    solver = "gurobi",
    PoolSearchMode = 2,
    PoolSolutions = 50,
    verbose = TRUE
  )
)
```

```{r}
pool_results <- extract_phd_pool_results(
  result_pool = result_pool,
  preference = preference,
  k = 50
)
```

```{r}
pool_results[[2]]$phd_assignment
```

```{r}
pool_results[[2]]$ta_detail
```

```{r}
# top 3 unique
n_plot <- 3

for (i in seq_len(n_plot)) {
  cur <- pool_results[[i]]

  p <- plot_phd_year_distribution(
    current_alloc_df = cur$phd_assignment,
    t1 = df_list$t1,
    g1 = df_list$g1,
    facet_by_seniority = TRUE,
    seniority_col = "seniority",
    facet_ncol = 4
  ) +
    ggplot2::labs(
      subtitle = paste0(
        "rank=", i,
        " | obj=", round(cur$objective_value, 4),
        " | TA min=", round(cur$ta_min, 4),
        " | TA max=", round(cur$ta_max, 4)
      )
    )

  print(p)
}
```

## Objective Comparison with manual output

```{r}
manual_output <- manual_output_current %>%
  filter(!is.na(Name), Name != "") %>%
  mutate(name_std = standardise_name(Name))

pref_lookup <- preference %>%
  transmute(
    i = row_number(),
    student_id = `S/No.`,
    student_name = Name,
    name_std = standardise_name(Name)
  )

ta_cols <- names(manual_output)[str_detect(names(manual_output), "-t$")]
gr_cols <- names(manual_output)[str_detect(names(manual_output), "-g$")]

manual_alloc <- pref_lookup %>%
  left_join(
    manual_output %>%
      transmute(
        name_std,
        manual_TA = rowSums(across(all_of(ta_cols)), na.rm = TRUE),
        manual_GR = rowSums(across(all_of(gr_cols)), na.rm = TRUE)
      ) %>%
      distinct(name_std, .keep_all = TRUE),
    by = "name_std"
  ) %>%
  transmute(
    student_id,
    student_name,
    manual_TA = replace_na(as.numeric(manual_TA), 0),
    manual_GR = replace_na(as.numeric(manual_GR), 0),
    manual_E = pmax(0, 4 - manual_TA - manual_GR)
  )
```

```{r}
# Must match the settings used in prepare_phd_model above.
obj_cfg <- list(alpha = 2, beta = 1, phi = 1, rho = 10, t_max_y1 = 1)
```

```{r}
# Solver objective for the model solution.
model_solver_obj <- as.numeric(result_0$objective_value)
if (length(model_solver_obj) == 0 || is.na(model_solver_obj[[1]])) {
  model_solver_obj <- as.numeric(result_0$additional_solver_output$ROI$message$objval)
}
model_solver_obj <- model_solver_obj[[1]]
```

```{r}
# Recalculate manual objective using manual TA by course + derived E.
manual_ta_detail <- manual_output %>%
  inner_join(pref_lookup, by = "name_std") %>%
  select(i, student_id, student_name, matches("-t$")) %>%
  pivot_longer(
    cols = matches("-t$"),
    names_to = "course_col",
    values_to = "ta_units"
  ) %>%
  mutate(
    ta_units = replace_na(as.numeric(ta_units), 0),
    course_code = norm_code(str_remove(course_col, "-t$")),
    j = match(course_code, courses)
  ) %>%
  filter(ta_units > 1e-8)

manual_ta_pref_sum <- manual_ta_detail %>%
  filter(!is.na(j)) %>%
  mutate(pref_score = P[cbind(i, j)]) %>%
  summarise(val = sum(pref_score * ta_units, na.rm = TRUE)) %>%
  pull(val)

manual_totals_by_i <- pref_lookup %>%
  select(i, student_id, student_name) %>%
  left_join(manual_alloc, by = c("student_id", "student_name")) %>%
  mutate(
    manual_TA = replace_na(as.numeric(manual_TA), 0),
    manual_E = replace_na(as.numeric(manual_E), 0)
  ) %>%
  arrange(i)
```

```{r}
manual_obj_breakdown <- compute_obj_from_components(
  ta_by_student = manual_totals_by_i$manual_TA,
  e_by_student = manual_totals_by_i$manual_E,
  ta_pref_sum = manual_ta_pref_sum,
  t1 = t1,
  s = s,
  t_max_y1 = obj_cfg$t_max_y1,
  alpha = obj_cfg$alpha,
  beta = obj_cfg$beta,
  phi = obj_cfg$phi,
  rho = obj_cfg$rho
)

manual_obj_breakdown
```

```{r}
model_solver_obj
```

## Manual-Style Solver Output

```{r}
job_output <- assign_job(
  model_result = result_0,
  student_df = preference,
  course_codes = courses,
  name_col = "Name"
)

job_output
```

```{r}
# check per-student TA / GR sessions from job_output
ta_cols_job <- names(job_output)[str_detect(names(job_output), "-t$")]
gr_cols_job <- names(job_output)[str_detect(names(job_output), "-g$")]
e_cols_job <- names(job_output)[str_detect(names(job_output), "-e$")]

job_output_summary <- job_output %>%
  transmute(
    Name,
    TA_sessions = rowSums(across(all_of(ta_cols_job)), na.rm = TRUE),
    GR_sessions = rowSums(across(all_of(gr_cols_job)), na.rm = TRUE),
    E_sessions = rowSums(across(all_of(e_cols_job)), na.rm = TRUE)
  ) %>%
  arrange(Name)

job_output_summary
```

## AY2420 Objective-Term Decomposition

```{r}
# Break down manual-vs-model objective gap into weighted terms.

model_ta_detail <- get_solution(result_0, X[i, j, r]) %>%
  filter(r == 1, value > 1e-8) %>%
  transmute(i = as.integer(i), j = as.integer(j), ta_units = as.numeric(value))

model_ta_pref_sum <- model_ta_detail %>%
  mutate(pref_score = P[cbind(i, j)]) %>%
  summarise(val = sum(pref_score * ta_units, na.rm = TRUE)) %>%
  pull(val)

model_totals_by_i <- pref_lookup %>%
  select(i, student_id, student_name) %>%
  left_join(
    alloc_df_0 %>%
      transmute(
        student_id,
        model_TA = as.numeric(TA),
        model_E = as.numeric(E)
      ),
    by = "student_id"
  ) %>%
  mutate(
    model_TA = replace_na(model_TA, 0),
    model_E = replace_na(model_E, 0)
  ) %>%
  arrange(i)

model_obj_breakdown <- compute_obj_from_components(
  ta_by_student = model_totals_by_i$model_TA,
  e_by_student = model_totals_by_i$model_E,
  ta_pref_sum = model_ta_pref_sum,
  t1 = t1,
  s = s,
  t_max_y1 = obj_cfg$t_max_y1,
  alpha = obj_cfg$alpha,
  beta = obj_cfg$beta,
  phi = obj_cfg$phi,
  rho = obj_cfg$rho
)

obj_term_breakdown <- bind_rows(
  Model = model_obj_breakdown,
  Manual = manual_obj_breakdown,
  .id = "Schedule"
) %>%
  mutate(
    fairness_term = obj_cfg$alpha * ta_spread,
    preference_term = -obj_cfg$beta * ta_pref_sum,
    seniority_e_term = -obj_cfg$phi * seniority_e_sum,
    y1_slack_term = obj_cfg$rho * y1_slack_sum
  ) %>%
  select(
    Schedule,
    fairness_term,
    preference_term,
    seniority_e_term,
    y1_slack_term,
    objective_recalc,
    ta_spread,
    ta_pref_sum,
    seniority_e_sum,
    y1_slack_sum
  )

obj_term_gap <- obj_term_breakdown %>%
  select(Schedule, fairness_term, preference_term, seniority_e_term, y1_slack_term, objective_recalc) %>%
  pivot_longer(-Schedule, names_to = "term", values_to = "value") %>%
  pivot_wider(names_from = Schedule, values_from = value) %>%
  mutate(
    manual_minus_model = Manual - Model,
    abs_gap = abs(manual_minus_model)
  ) %>%
  arrange(desc(abs_gap))

model_obj_check <- tibble(
  solver_objective = model_solver_obj,
  recalc_objective = obj_term_breakdown %>% filter(Schedule == "Model") %>% pull(objective_recalc),
  recalc_minus_solver = recalc_objective - solver_objective
)

obj_term_gap
model_obj_check
```

```{r}
# Decomposition chart: weighted objective-term differences (Manual - Model).
term_plot_df <- obj_term_gap %>%
  filter(term %in% c("preference_term", "fairness_term", "seniority_e_term", "y1_slack_term")) %>%
  mutate(
    term = recode(
      term,
      preference_term = "Preference",
      fairness_term = "Fairness",
      seniority_e_term = "Seniority-E",
      y1_slack_term = "Year-1 Slack"
    ),
    term = factor(term, levels = c("Preference", "Fairness", "Seniority-E", "Year-1 Slack")),
    direction = if_else(manual_minus_model >= 0, "Higher in Manual", "Lower in Manual"),
    label_hjust = if_else(manual_minus_model >= 0, -0.15, 1.15)
  )

p_obj_gap <- ggplot(term_plot_df, aes(x = term, y = manual_minus_model, fill = direction)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey45") +
  geom_col(width = 0.62) +
  geom_text(aes(label = round(manual_minus_model, 1), hjust = label_hjust), size = 3) +
  coord_flip() +
  scale_fill_manual(values = c("Higher in Manual" = "#e61b00", "Lower in Manual" = "#0072B2")) +
  labs(x = NULL, y = "Weighted term difference (Manual - Model)", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

p_obj_gap
```

```{r}
# Side-by-side TA preference composition (units): model vs manual.
pref_levels <- c("First", "Second", "Third", "Unranked")

map_pref_rank <- function(score) {
  case_when(
    score >= 3 ~ "First",
    score == 2 ~ "Second",
    score == 1 ~ "Third",
    TRUE ~ "Unranked"
  )
}

model_pref_units <- model_ta_detail %>%
  mutate(
    pref_score = P[cbind(i, j)],
    preference = map_pref_rank(pref_score)
  ) %>%
  group_by(preference) %>%
  summarise(units = sum(ta_units, na.rm = TRUE), .groups = "drop")

manual_pref_units <- manual_ta_detail %>%
  mutate(
    pref_score = if_else(!is.na(j), P[cbind(i, j)], -99),
    preference = map_pref_rank(pref_score)
  ) %>%
  group_by(preference) %>%
  summarise(units = sum(ta_units, na.rm = TRUE), .groups = "drop")

pref_plot_df <- bind_rows(
  model_pref_units %>% mutate(Schedule = "Model"),
  manual_pref_units %>% mutate(Schedule = "Manual")
) %>%
  mutate(
    preference = factor(preference, levels = pref_levels),
    Schedule = factor(Schedule, levels = c("Model", "Manual"))
  ) %>%
  tidyr::complete(Schedule, preference, fill = list(units = 0))

p_pref_side <- ggplot(pref_plot_df, aes(x = preference, y = units, fill = Schedule)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.62) +
  geom_text(
    aes(label = round(units, 1)),
    position = position_dodge(width = 0.72),
    vjust = -0.25,
    size = 3
  ) +
  scale_fill_manual(values = c("Model" = "#0072B2", "Manual" = "#e61b00")) +
  labs(x = NULL, y = "TA units", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

p_pref_side
```
